---
title: "Jensenii-Mulieris Analysis"
author: "Erin Yuan"
date: "2022-10-25"
output: html_document
---

This Rmd file takes raw counts of SGBs, converts counts to proportions using a correction factor for scaling, and visualizes them using stacked bar plots. 

```{r setup, warning = F}
# install packages
# install.packages("ggplot2")
# install.packages("reshape2")
# install.packages("ape")
# install.packages("dplyr")
# install.packages("vegan")
# install.packages("pheatmap")
# install.packages("cowplot")
# install.packages("devtools")
# devtools::install_github("jaredhuling/jcolors")
# install.packages("viridis)

# load packages
library(ggplot2)
library(reshape2)
library(ape)
library(dplyr)
library(vegan)
library(pheatmap)
library(cowplot)
library(jcolors)
library(viridis)
```

# Data cleaning
```{r}
# load data
counts <- read.table("data/SGB_read_counts_NorthAmerican_26July2021.tab", sep = "\t", check.names = T, header = T, row.names = 1) # rows = samples, cols = SGBs
correction <- read.table("data/correction_factor_8July2021.tab", sep = "\t", check.names = T, header = F, row.names = NULL) # rows = SGB names, col = correction factor 
metadata <- read.table("data/combined_metadata_with_VIRGO_5_Nov_2020.tab", sep = "\t", check.names = T, header = T, row.names = 1) # rows = samples, cols = age, ethnicity, status, country, study 

## preprocessing for counts
# shift the column names to the left 
counts_fixed = counts[,-c(337)] # 337 = total number of columns  
counts_fixed = counts_fixed[,c(-1)]
colnames(counts_fixed) = colnames(counts)[seq(3,length(counts[1,]))]
counts_fixed <- counts_fixed[, 1:335] # rows = samples, cols = SGBs

# delete row with sample CM_G1-2_MERGE because there is no metadata associated with it  
delete_row = which(rownames(counts_fixed) == "CM_G1-2_MERGE")
counts_fixed = counts_fixed[c(-140), ]

## metadata
# delete the first row because it is empty and only includes the column header for SGBs
metadata <- metadata[-1, ]

# remove NAs for sample CM_G1-2_MERGE
# metadata <- metadata %>%
#   filter(!is.na(study))
```

Scaling SGB counts: make scaled counts using correction factor
```{r}
# merge correction with counts_fixed
counts_t <- as.data.frame(t(counts_fixed)) # rows = SGBs, columns = samples 
counts_t <- counts_t %>%
  mutate(sgbs = rownames(counts_t))
counts_correction_t <- left_join(counts_t, correction, by = c("sgbs" = "V1"))

# remove SGBs with NA correction factor
counts_correction_noNA_t <- counts_correction_t %>%
  filter(!is.na(V2))
print(paste0("# of SGBs removed bc NA correction factor: ", nrow(counts_correction_t) - nrow(counts_correction_noNA_t)))

# multiply each row's entry by the correction factor 
counts_corrected_t <- counts_correction_noNA_t
for (row in 1:nrow(counts_corrected_t)) { # iterate over each row (sgb)
  for (col in 1:(ncol(counts_corrected_t)-2)) {
    counts_corrected_t[row, col] = counts_corrected_t[row, col] * counts_corrected_t[row, ncol(counts_corrected_t)]
  }
}

# rename rownames to be sgbs
sgbs <- counts_corrected_t[, ncol(counts_corrected_t)-1]  
rownames(counts_corrected_t) = sgbs
counts_corrected_t <- counts_corrected_t[, 1:(ncol(counts_corrected_t)-2)]
counts_corrected <- as.data.frame(t(counts_corrected_t)) # rows = samples, cols = SGBs

# find the total for a sample
counts_total <- counts_corrected %>%
  mutate(total = rowSums(counts_corrected))

# remove 0s
counts_nz_total <- counts_total[-which(rowSums(counts_total) == 0), ] # if sample (column) sums to 0, then remove

# find proportions
counts_prop <- counts_nz_total
for (row in 1:nrow(counts_prop)) { # iterate over each row (sample)
  for (col in 1:(ncol(counts_prop)-1)) {
    counts_prop[row, col] = as.numeric(counts_prop[row, col] / counts_prop[row, ncol(counts_prop)])
  }
}
counts_prop <- counts_prop[, 1:ncol(counts_prop)-1] # rows = samples, cols = SGBs

# make correlation matrix 
corr <- cor(counts_prop, method = "spearman")
corr_jensenii <- corr[which(rownames(corr) %in% c("SGB14", "SGB156")), ]
pheatmap(corr_jensenii)
```

Cutoff 
```{r}
# find SGBs that have correlation more than [cutoff]
cutoff = 0.2
above_cutoff = c() # create new vector to store names of SGBs that have correlation more than [cutoff]
for (row in c(1, 2)) { 
  for (col in 1:ncol(corr_jensenii)) {
    if (corr_jensenii[row, col] > cutoff) {
      above_cutoff <- c(above_cutoff, colnames(corr_jensenii)[col])
    }
  }
}
corr_jensenii_above_cutoff <- corr_jensenii[, which(colnames(corr_jensenii) %in% above_cutoff)]
pheatmap(corr_jensenii_above_cutoff)
```

# Figure 2a: Stacked bars for samples vs. abundances of jensenii, not divided by study
```{r}
# find top [max] most abundant SGBs (sum of abundances is the highest)
max = 10
counts_prop_t <- as.data.frame(t(counts_prop)) 
counts_prop_sum <- counts_prop_t %>%
  mutate(sgbSum = rowSums(counts_prop_t)) %>% # create new col with row (sgb) sums across all samples 
  arrange(desc(sgbSum)) %>% # arrange from highest to lowest sums
  select(sgbSum) # view only sums
top <- rownames(counts_prop_sum)[1:max] # select only top [max]
# test: these two lines should match
# top 
# head(counts_prop_sum, n = 10)

# limit to only samples that have SGB14 or SGB156 (jensenii)
onlyJenseniiCutoff = 0.01 # remove samples < 1% of any jensenii
counts_prop_jensenii <- counts_prop %>%
  select(SGB14, SGB156)
counts_prop_jensenii_above_cutoff <- counts_prop_jensenii %>%
  mutate(sampleSumJensenii = rowSums(counts_prop_jensenii)) %>%
  filter(sampleSumJensenii > onlyJenseniiCutoff)
print(paste0("# of samples removed bc no jensenii: ", nrow(counts_prop_jensenii) - nrow(counts_prop_jensenii_above_cutoff))) 
samplesWithJensenii <- rownames(counts_prop_jensenii_above_cutoff)
counts_prop_samplesWithJensenii <- counts_prop %>%
  mutate(samples = rownames(counts_prop)) %>%
  filter(samples %in% samplesWithJensenii)
print(paste0("# of samples removed bc no jensenii: ", nrow(counts_prop) - nrow(counts_prop_samplesWithJensenii))) 
```

ggplot
```{r}
# melt 
counts_prop_melt <- melt(counts_prop_samplesWithJensenii, id.vars = "samples", variable.name = "sgb", value.name = "percentage")

# create new label for top10 and SGB14 and SGB156 (jensenii) 
counts_prop_jensenii_melt <- counts_prop_melt %>%
  mutate(label = ifelse(sgb %in% c("SGB14", "SGB156", top), as.character(sgb), "Other"))

# create a factor for label
counts_prop_jensenii_melt$label <- factor(counts_prop_jensenii_melt$label, levels = c("SGB156", "SGB14", "SGB164", "SGB58", "SGB59", "SGB118", "SGB91", "SGB359", "SGB227", "SGB144", "Other"))

# reorder the x-axis (samples) 
counts_prop_jensenii_SGB14 <- counts_prop_jensenii_melt %>% # make new data frame of only SGB14
  filter(sgb == "SGB14") %>%
  select(-sgb)
counts_prop_jensenii_order <- counts_prop_jensenii_above_cutoff %>%
  mutate(SGB156_ordered = ifelse(SGB14 > SGB156, 0, as.numeric(SGB156))) %>%
  select(SGB156_ordered)
counts_prop_jensenii_melt$samples <- factor(counts_prop_jensenii_melt$samples, levels = counts_prop_jensenii_melt$samples[order(-counts_prop_jensenii_order$SGB156_ordered, counts_prop_jensenii_SGB14$percentage)]) # ordered by SGB156 then SGB14

# plot
n_abundance = nrow(counts_prop_jensenii_above_cutoff)
sgb_abundances = ggplot(data = counts_prop_jensenii_melt, mapping = aes(x = samples, y = percentage)) +
  geom_col(aes(fill = label)) + 
  labs( # title = "Jensenii Figure 2a Top", 
       x = "Samples", 
       y = "Relative Abundance", 
       # caption = paste0("n = ", n_abundance)
       ) +  
  theme_minimal() + 
  theme(legend.position = "none", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank()) +
  scale_fill_jcolors(palette = "pal8")
sgb_abundances
# ggsave("figures/jensenii_figure2a_top.pdf")
```

# Figure 2a strips
Stacked bars for samples vs. abundances for jensenii
```{r}
# recalculate proportions for only SGB14 and SGB156 (jensenii)
counts_prop_jensenii_only <- counts_prop_jensenii_above_cutoff
for (row in 1:nrow(counts_prop_jensenii_only)) { # iterate over each row (sample)
  for (col in 1:(ncol(counts_prop_jensenii_only)-1)) { 
    counts_prop_jensenii_only[row, col] = as.numeric(counts_prop_jensenii_only[row, col] / counts_prop_jensenii_only[row, ncol(counts_prop_jensenii_only)])
  }
}
counts_prop_jensenii_only <- counts_prop_jensenii_only[, 1:ncol(counts_prop_jensenii_only)-1] # rows = samples, cols = SGBs 
counts_prop_jensenii_only_samples <- counts_prop_jensenii_only %>%
  mutate(samples = rownames(counts_prop_jensenii_only))
```

ggplot
```{r}
# melt 
counts_prop_jensenii_only_melt <- melt(counts_prop_jensenii_only_samples, id.vars = "samples", variable.name = "sgb", value.name = "percentage")

# change order of factor for sgb
counts_prop_jensenii_only_melt$sgb <- factor(counts_prop_jensenii_only_melt$sgb, levels = c("SGB156", "SGB14"))

# reorder by the SAME ORDER as the above plot
counts_prop_jensenii_only_melt$samples <- factor(counts_prop_jensenii_only_melt$samples, levels = counts_prop_jensenii_only_melt$samples[order(-counts_prop_jensenii_order$SGB156_ordered, counts_prop_jensenii_SGB14$percentage)]) # ordered by SGB156 then SGB14

# plot
jensenii_ratios = ggplot(data = counts_prop_jensenii_only_melt, mapping = aes(x = samples, y = percentage)) +
  geom_col(aes(fill = sgb)) + 
  labs( # title = "Jensenii Figure 2a Ratio", 
       x = "Samples", 
       y = "Relative Abundance", 
       # caption = paste0("n = ", n_abundance)
       ) +  
  scale_fill_jcolors(palette = "pal8") + 
  theme_minimal() + 
  theme(legend.position = "none", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.y=element_blank()) 
jensenii_ratios
# ggsave("figures/jensenii_figure2a_ratio_legend.pdf")
```

```{r}
# find number of 100% SGB14, 100% SGB156, and both
counts_prop_jensenii_allSGB14 <- counts_prop_jensenii_only %>%
  filter(SGB14 > 0.99)
print(paste0("# of samples with only L. jensenii: ", nrow(counts_prop_jensenii_allSGB14)))
counts_prop_jensenii_allSGB156 <- counts_prop_jensenii_only %>%
  filter(SGB156 > 0.99)
print(paste0("# of samples with only L. mulieris: ", nrow(counts_prop_jensenii_allSGB156)))
print(paste0("# of samples with both L. jensenii and L. mulieris: ", nrow(counts_prop_jensenii_only)-nrow(counts_prop_jensenii_allSGB14)-nrow(counts_prop_jensenii_allSGB156)))
```

Metadata strips
```{r}
metadata_samples <- metadata %>%
  mutate(samples = rownames(metadata)) %>% # add sample names as a column
  mutate(stat = 1) %>% # add column called stat = 1
  select(-country, -age) # remove country column bc all from NAm, remove age bc not relevant
counts_prop_jensenii_above_cutoff_samples <- counts_prop_jensenii_above_cutoff %>%
  mutate(samples = rownames(counts_prop_jensenii_above_cutoff))

# merge metadata with proportions of jensenii only
samples_with_metadata <- left_join(counts_prop_jensenii_above_cutoff_samples, metadata_samples) %>%
  select(-SGB14, -SGB156, -sampleSumJensenii) %>%
  mutate(samples = rownames(counts_prop_jensenii_above_cutoff_samples))
```

1. by study
```{r}
strip_study_data <- samples_with_metadata %>%
  select(samples, study, stat)

# reorder 
strip_study_data$samples <- factor(strip_study_data$samples, levels = strip_study_data$samples[order(-counts_prop_jensenii_order$SGB156_ordered, counts_prop_jensenii_SGB14$percentage)])

# plot
strip_study_plot = ggplot(data = strip_study_data, aes(x = samples, y = stat, fill = study)) + 
  geom_tile(colour="white",size=0.25) + 
  labs(# title = "Study", 
       # caption = paste0("n = ", n_abundance)
      ) + 
  theme(legend.position = "none", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),  
        axis.title.x=element_blank(),  
        axis.title.y= element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank(), 
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        text = element_text(size=9)) + 
  scale_fill_viridis(discrete = TRUE, option = "C")
strip_study_plot
# ggsave("figures/jensenii_figure2a_study_legend.pdf")
```

2. by reproductive status
```{r}
strip_reprod_data <- samples_with_metadata %>%
  select(samples, status, stat)

# reorder 
strip_reprod_data$samples <- factor(strip_reprod_data$samples, levels = strip_reprod_data$samples[order(-counts_prop_jensenii_order$SGB156_ordered, counts_prop_jensenii_SGB14$percentage)])

# plot
strip_reprod_plot = ggplot(data = strip_reprod_data, aes(x = samples, y = stat, fill = status)) + 
  geom_tile(colour="white",size=0.25) + 
  labs(# title = "Reproductive Status", 
       # caption = paste0("n = ", n_abundance)
      ) + 
  theme(legend.position = "none", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),  
        axis.title.x=element_blank(),  
        axis.title.y= element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank(), 
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        text = element_text(size=9)) + 
  scale_fill_viridis(discrete = TRUE)
strip_reprod_plot
# ggsave("figures/jensenii_figure2a_status_legend.pdf")
```

3. by ethnicity
```{r}
strip_ethnicity_data <- samples_with_metadata %>%
  select(samples, ethnicity, stat)

# reorder 
strip_ethnicity_data$samples <- factor(strip_ethnicity_data$samples, levels = strip_ethnicity_data$samples[order(-counts_prop_jensenii_order$SGB156_ordered, counts_prop_jensenii_SGB14$percentage)])

# plot
cols <- c("AA" = "red", "Pac" = "blue", "W" = "darkgreen", "na" = "gray")
strip_ethnicity_plot = ggplot(data = strip_ethnicity_data, aes(x = samples, y = stat, fill = ethnicity)) + 
  geom_tile(colour="white",size=0.25) + 
  labs(# title = "Ethnicity", 
       # caption = paste0("n = ", n_abundance)
       ) + 
  theme(legend.position = "none", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),  
        axis.title.x=element_blank(),  
        axis.title.y= element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank(), 
        axis.line=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        text = element_text(size=9)) +
  scale_fill_manual(values = cols)
strip_ethnicity_plot
# ggsave("figures/jensenii_figure2a_ethnicity_legend.pdf")
```

Cowplot
```{r, fig.width = 5}
plot_grid(sgb_abundances, jensenii_ratios, strip_study_plot, strip_reprod_plot, strip_ethnicity_plot, align = "v", ncol = 1, rel_heights = c(7, 1, 0.5, 0.5, 0.5))
ggsave("figures/jensenii_figure2a.pdf")
```

# Supplemental Figure X
```{r}
# create new dataframe with presence/absence information
presentDefinition = 0.01
counts_prop_jensenii_presence <- counts_prop_jensenii %>% 
  mutate(SGB14_present = ifelse(SGB14 > presentDefinition, "Yes", "No")) %>%
  mutate(SGB156_present = ifelse(SGB156 > presentDefinition, "Yes", "No")) %>%
  select(SGB14_present, SGB156_present) %>%
  mutate(samples = rownames(counts_prop_jensenii))

# merge with metadata
presence_metadata <- left_join(counts_prop_jensenii_presence, metadata_samples)
```

```{r}
# create 3 new dataframes for each feature

# study
presence_metadata_study_SGB14 <- presence_metadata %>%
  select(samples, SGB14_present, study, -stat) %>%
  group_by(study, SGB14_present) %>%
  summarise(counts = n()) %>%
  mutate(prop = counts / sum(counts)) %>%
  filter(SGB14_present == "Yes") %>%
  rename(sgb = SGB14_present) %>%
  mutate(sgb = recode(sgb, "Yes" = "SGB14"))
presence_metadata_study_SGB156 <- presence_metadata %>%
  select(samples, SGB156_present, study, -stat) %>%
  group_by(study, SGB156_present) %>%
  summarise(counts = n()) %>%
  mutate(prop = counts / sum(counts)) %>%
  filter(SGB156_present == "Yes") %>%
  rename(sgb = SGB156_present) %>%
  mutate(sgb = recode(sgb, "Yes" = "SGB156"))
presence_metadata_study <- rbind (presence_metadata_study_SGB14, presence_metadata_study_SGB156)
presence_metadata_study$sgb <- factor(presence_metadata_study$sgb, levels = c("SGB14", "SGB156")) # change order of factor for sgb

# reproductive status
presence_metadata_reprod_SGB14 <- presence_metadata %>%
  select(samples, SGB14_present, status, -stat) %>%
  group_by(status, SGB14_present) %>%
  summarise(counts = n()) %>%
  mutate(prop = counts / sum(counts)) %>%
  filter(SGB14_present == "Yes") %>%
  rename(sgb = SGB14_present) %>%
  mutate(sgb = recode(sgb, "Yes" = "SGB14"))
presence_metadata_reprod_SGB156 <- presence_metadata %>%
  select(samples, SGB156_present, status, -stat) %>%
  group_by(status, SGB156_present) %>%
  summarise(counts = n()) %>%
  mutate(prop = counts / sum(counts)) %>%
  filter(SGB156_present == "Yes") %>%
  rename(sgb = SGB156_present) %>%
  mutate(sgb = recode(sgb, "Yes" = "SGB156"))
presence_metadata_reprod <- rbind (presence_metadata_reprod_SGB14, presence_metadata_reprod_SGB156)
presence_metadata_reprod$sgb <- factor(presence_metadata_reprod$sgb, levels = c("SGB14", "SGB156")) # change order of factor for sgb

# ethnicity
presence_metadata_ethnicity_SGB14 <- presence_metadata %>%
  select(samples, SGB14_present, ethnicity, -stat) %>%
  group_by(ethnicity, SGB14_present) %>%
  summarise(counts = n()) %>% 
  ungroup() %>%
  add_row(ethnicity = "As", SGB14_present = "Yes", counts = 0) %>%
  add_row(ethnicity = "H", SGB14_present = "Yes", counts = 0) %>%
  add_row(ethnicity = "Other", SGB14_present = "Yes", counts = 0) %>%
  add_row(ethnicity = "Pac", SGB14_present = "Yes", counts = 0) %>%
  group_by(ethnicity) %>%
  mutate(prop = counts / sum(counts)) %>%
  filter(SGB14_present == "Yes") %>%
  rename(sgb = SGB14_present) %>%
  mutate(sgb = recode(sgb, "Yes" = "SGB14"))
presence_metadata_ethnicity_SGB156 <- presence_metadata %>%
  select(samples, SGB156_present, ethnicity, -stat) %>%
  group_by(ethnicity, SGB156_present) %>%
  summarise(counts = n()) %>%
  ungroup() %>%
  add_row(ethnicity = "As", SGB156_present = "Yes", counts = 0) %>%
  add_row(ethnicity = "H", SGB156_present = "Yes", counts = 0) %>%
  add_row(ethnicity = "Other", SGB156_present = "Yes", counts = 0) %>%
  group_by(ethnicity) %>%
  mutate(prop = counts / sum(counts)) %>%
  filter(SGB156_present == "Yes") %>%
  rename(sgb = SGB156_present) %>%
  mutate(sgb = recode(sgb, "Yes" = "SGB156"))
presence_metadata_ethnicity <- rbind (presence_metadata_ethnicity_SGB14, presence_metadata_ethnicity_SGB156)
presence_metadata_ethnicity$sgb <- factor(presence_metadata_ethnicity$sgb, levels = c("SGB14", "SGB156")) # change order of factor for sgb

# set color limits for heatmap
rng = range(presence_metadata_study$prop, presence_metadata_reprod$prop, presence_metadata_ethnicity$prop)
```

ggplots
1. by study
```{r}
# plot
ggplot(data = presence_metadata_study, aes(x = study, y = sgb, fill = prop)) + 
  geom_tile(colour="black",size=0.25) + 
  labs(# title = "Study", 
       # caption = paste0("n = ", n)
       ) + 
  theme(# legend.position = "none", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),  
        # axis.title.x=element_blank(),  
        axis.title.y= element_blank(), 
        axis.line=element_blank(),
        # axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        text = element_text(size=9)) + 
  scale_fill_gradient2(# low="purple", high="yellow", 
               # midpoint=mean(rng),    
               # breaks=seq(0,1,0.25), #breaks in the scale bar
               limits=c(floor(rng[1]), ceiling(rng[2])))
ggsave("figures/jensenii_supplemental_figure_study_legend.pdf")
```

2. by reproductive status 
```{r}
ggplot(data = presence_metadata_reprod, aes(x = status, y = sgb, fill = prop)) + 
  geom_tile(colour="black",size=0.25) + 
  labs(# title = "Reproductive Status", 
       # caption = paste0("n = ", n)
       ) + 
  theme(# legend.position = "none", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),  
        # axis.title.x=element_blank(),  
        axis.title.y= element_blank(), 
        axis.line=element_blank(),
        # axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        text = element_text(size=9)) +
   scale_fill_gradient2(# low="purple", high="yellow", 
               # midpoint=mean(rng),    
               # breaks=seq(0,1,0.25), #breaks in the scale bar
               limits=c(floor(rng[1]), ceiling(rng[2])))
ggsave("figures/jensenii_supplemental_figure_reprod_legend.pdf")
```

3. by ethnicity
```{r}
ggplot(data = presence_metadata_ethnicity, aes(x = ethnicity, y = sgb, fill = prop)) + 
  geom_tile(colour="black",size=0.25) + 
  labs(# title = "Ethnicity", 
       # caption = paste0("n = ", n)
       ) + 
  theme(# legend.position = "none", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),  
        # axis.title.x=element_blank(),  
        axis.title.y= element_blank(), 
        axis.line=element_blank(),
        # axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        text = element_text(size=9)) + 
   scale_fill_gradient2(# low="purple", high="yellow", 
               # midpoint=mean(rng),    
               # breaks=seq(0,1,0.25), #breaks in the scale bar
               limits=c(floor(rng[1]), ceiling(rng[2])))
ggsave("figures/jensenii_supplemental_figure_ethnicity_legend.pdf")
```

# Extra: Stacked bars divided by study 
```{r}
# make 3 separate dataframes for 3 separate studies
counts_metadata_hmp_melt <- counts_metadata_jensenii_melt %>%
  filter(study == "HMP")
counts_metadata_msflash_melt <- counts_metadata_jensenii_melt %>%
  filter(study == "MsFlash") 
counts_metadata_relman_melt <- counts_metadata_jensenii_melt %>%
  filter(study == "Relman")
```

ggplots
1. HMP
```{r}
# find number of samples
counts_metadata_hmp_melt_unique <- counts_metadata_hmp_melt %>%
  select(samples) %>%
  distinct()
n_hmp = nrow(counts_metadata_hmp_melt_unique)

# reorder the x-axis (samples) 
counts_metadata_hmp_melt_SGB14 <- counts_metadata_hmp_melt %>% # make new data frame of only SGB14
  filter(sgb == "SGB14") %>%
  select(-sgb)
counts_metadata_hmp_melt_SGB156 <- counts_metadata_hmp_melt %>% # make new data frame of only SGB156
  filter(sgb == "SGB156") %>%
  select(-sgb)
counts_metadata_hmp_melt$samples <- factor(counts_metadata_hmp_melt$samples, levels = counts_metadata_hmp_melt$samples[order(-counts_metadata_hmp_melt_SGB14$percentage, counts_metadata_hmp_melt_SGB156$percentage)])

# plot
ggplot(data = counts_metadata_hmp_melt, mapping = aes(x = samples, y = percentage)) +
  geom_col(aes(fill = label)) + 
  labs(title = "Jensenii Figure 2b HMP", 
       x = "Samples", 
       y = "Relative Abundance", 
       caption = paste0("n = ", n_hmp)) +  
  scale_fill_discrete(name = "SGBs") + 
  theme_minimal() + 
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) 
# ggsave("jensenii_figure2b_hmp.pdf")
```

2. MsFlash
```{r}
# find number of samples
counts_metadata_msflash_melt_unique <- counts_metadata_msflash_melt %>%
  select(samples) %>%
  distinct()
n_msflash = nrow(counts_metadata_msflash_melt_unique)

# reorder the x-axis (samples) 
counts_metadata_msflash_melt_SGB14 <- counts_metadata_msflash_melt %>% # make new data frame of only SGB14
  filter(sgb == "SGB14") %>%
  select(-sgb)
counts_metadata_msflash_melt_SGB156 <- counts_metadata_msflash_melt %>% # make new data frame of only SGB156
  filter(sgb == "SGB156") %>%
  select(-sgb)
counts_metadata_msflash_melt$samples <- factor(counts_metadata_msflash_melt$samples, levels = counts_metadata_msflash_melt$samples[order(-counts_metadata_msflash_melt_SGB14$percentage, counts_metadata_msflash_melt_SGB156$percentage)])

# plot
ggplot(data = counts_metadata_msflash_melt, mapping = aes(x = samples, y = percentage)) +
  geom_col(aes(fill = label)) + 
  labs(title = "Jensenii Figure 2b MsFlash", 
       x = "Samples", 
       y = "Relative Abundance", 
       caption = paste0("n = ", n_msflash)) +  
  scale_fill_discrete(name = "SGBs") + 
  theme_minimal() + 
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) 
# ggsave("jensenii_figure2b_msflash.pdf")
```

3. Relman
```{r}
# find number of samples
counts_metadata_relman_melt_unique <- counts_metadata_relman_melt %>%
  select(samples) %>%
  distinct()
n_relman = nrow(counts_metadata_relman_melt_unique)

# reorder the x-axis (samples) 
counts_metadata_relman_melt_SGB14 <- counts_metadata_relman_melt %>% # make new data frame of only SGB14
  filter(sgb == "SGB14") %>%
  select(-sgb)
counts_metadata_relman_melt_SGB156 <- counts_metadata_relman_melt %>% # make new data frame of only SGB156
  filter(sgb == "SGB156") %>%
  select(-sgb)
counts_metadata_relman_melt$samples <- factor(counts_metadata_relman_melt$samples, levels = counts_metadata_relman_melt$samples[order(-counts_metadata_relman_melt_SGB14$percentage, counts_metadata_relman_melt_SGB156$percentage)])

# plot
ggplot(data = counts_metadata_relman_melt, mapping = aes(x = samples, y = percentage)) +
  geom_col(aes(fill = label)) + 
  labs(title = "Jensenii Figure 2b Relman", 
       x = "Samples", 
       y = "Relative Abundance", 
       caption = paste0("n = ", n_relman)) +  
  scale_fill_discrete(name = "SGBs") + 
  theme_minimal() + 
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) 
# ggsave("jensenii_figure2b_relman.pdf")
```